Processing large surveys with package MRP-KIT

### 2021-07-11 case-study/Mapping_CCES_to_ACS.Rmd
checkpointing

The CCES and ACS samples record the education level of the respondant
in variables `educ` and `EDUC`, respectively.
Both the CCES and ACS record the highest level of education completed,
but the CCES records the degree obtained whereas the ACS records the year completed.
```{r}
levels(cces_2018_subset$educ)
levels(acs_2018_subset$EDUC)
```
We choose to align these categories so that the CCES label "2-year"
maps to the ACS labels for 2 and 3 years of college.
Mpping "Some College" to both 1 year and 3 years of college is also possible,
on the rationale that the goal was a 4-year college degree, and that it wasn't obtained.
The choice depends on both an understanding of the survey data and the particular goals
of your analysis.

CCES | ACS
---- | ---
"No HS" | "N/A or no schooling"
"No HS" | "Nursery school to grade 4"
"No HS" | "Grade 5, 6, 7, or 8"
"No HS" | "Grade 9"
"No HS" | "Grade 10"
"No HS" | "Grade 11"
"High School Graduate" | "Grade 12"
"Some College" | "1 year of college"
"2-year" | "2 years of college"
"2-year" | "3 years of college"
"4-year" | "4 years of college"
"Post-grad" | "5+ years of college"

```{r}
educ_mapping = list(
    "No HS" = "N/A or no schooling",
    "No HS" = "Nursery school to grade 4",
    "No HS" = "Grade 5, 6, 7, or 8",
    "No HS" = "Grade 9",
    "No HS" = "Grade 10",
    "No HS" = "Grade 11",
    "High School Graduate" = "Grade 12",
    "Some College" = "1 year of college",
    "2-year" = "2 years of college",
    "2-year" = "3 years of college",
    "4-year" = "4 years of college",
    "Post-grad" = "5+ years of college"
)
q_educ <- SurveyQuestion$new(name = "educ",
    col_names = c("educ", "EDUC"),
    values_map= educ_mapping)
q_educ$name()
q_educ$col_names()
q_educ$values()
```

@@

```{r}
# create column which conflates spouse present, absent
acs_marst2 <- acs_2018_subset$MARST
levels(acs_marst2) = c(levels(acs_marst2), "Married")
acs_marst2[acs_marst2 =="Married, spouse present"] <- "Married"
acs_marst2[acs_marst2 =="Married, spouse absent"] <- "Married"

# add to acs_2018_subset data.frame
acs_2018_subset$marst2 = acs_marst2
```
Now we can create a mapping which maps both "Married" and ""Domestic / civil partnership"
to the the (new) ACS label "Married".^[
Alternatively, we could try to access more ACS variables which account for
same sex marriages and civil unions and combine information across multiple columns, as we did
for ACS variable race2.]

```{r}
marstat_mapping = list(
    "Married" = "Married",
    "Separated" = "Separated",
    "Divorced" = "Divorced",
    "Widowed" = "Widowed",
    "Never married" = "Never married",
    "Domestic / civil partnership" = "Married"
)
q_marstat <- SurveyQuestion$new(name = "marstat",
    col_names = c("marstat", "marst2"),
    values_map= marstat_mapping)
q_marstat$name()
q_marstat$col_names()
q_marstat$values()
```

@@

A SurveyData object requires a dataframe, a question list, and a response list.
A question list is a list of tag-value pairs where the tag (name) is the dataframe column name and the value is a description of corresponding survey question.
For the CCES, the dataframe column descriptions are lists in the attrbute "var.labels".
As before, we create this list programatically using the `as.list(setNames(tags,values))` trick.
```{r, EVAL=FALSE}
cces_2018_questions <- as.list(setNames(attr(cces_2018_subset,"var.labels"), names(cces_2018_subset)))
head(cces_2018_questions)
```
For the ACS, we use function `ipumsr::ipums_var_label` to get a short description of the variable
from the metadata in the DDI object.
```{r, EVAL=FALSE}
acs_2018_var_labels <- c()
for (name in names(acs_2018_subset)) {
    acs_2018_var_labels <- c(acs_2018_var_labels, ipums_var_label(acs_2018_subset[[name]]))
}
acs_2018_questions <- as.list(setNames(acs_2018_var_labels, names(acs_2018_subset)))
head(acs_2018_questions)
```

A response list is a list of name-value pairs which map the column name to the set of
possible response values.
We can create this list using a for loop, or by using R's `apply` function.
```{r, EVAL=FALSE}
# create response list using a for loop
cces_2018_responses <- list()
for (name in names(cces_2018_subset)) {
    cces_2018_responses[[name]] <- levels(as.factor(cces_2018_subset[[name]]))
}
head(cces_2018_responses)

acs_2018_responses <- list()
for (name in names(acs_2018_subset)) {
    acs_2018_responses[[name]] <- levels(as_factor(acs_2018_subset[[name]]))
}
acs_2018_responses[["MARST"]]
acs_2018_responses[["race2"]]
acs_2018_responses[["birthyr_cohort"]]

# cces_2018_subset is a dataframe; R's `apply` function would also work
# cces_2018_responses <- apply(cces_2018_subset, 2, function(x) levels(as.factor(x)))
# if we coerce acs_2018_subset to dataframe, can do the same
# acs_2018_responses <- apply(as.data.frame(cces_2018_subset), 2, function(x) levels(as.factor(x)))
```

We use these lists to create SurveyData objects for both the CCES and ACS.
```{r, EVAL=FALSE}
cces_2018_data <- SurveyData$new(cces_2018_subset, cces_2018_questions, cces_2018_responses)
cces_2018_data
acs_2018_data <- SurveyData$new(acs_2018_subset, acs_2018_questions, acs_2018_responses)
acs_2018_data
```
Now we can create the mapping object from the CCES to the ACS.
```{r, EVAL=FALSE}
cces_acs_mapping <- SurveyMap$new(cces_2018_data, acs_2018_data, q_sex, q_state, q_race, q_educ)
# poststrat table size: 2 * 51 * 8 * 6
```

>>>>

The CCES and ACS samples record the education level of the respondant
in variables `educ` and `EDUC`, respectively.
Both the CCES and ACS record the highest level of education completed,
but the CCES records the degree obtained whereas the ACS records the year completed.
```{r}
levels(cces_2018_subset$educ)
levels(acs_2018_subset$EDUC)
```

For the ACS data, a possible label is "3 years of college" however
in the ACS subset used here, this label isn't used.
If it were used, we would have a choice between mapping it to CCES
labels "Some college" or "2-year", on the rational that if the
 goal was a 4-year college degree, and that it wasn't obtained.
The choice depends on both an understanding of the survey data and the particular goals
of your analysis.

CCES | ACS
---- | ---
"No HS" | "N/A or no schooling"
"No HS" | "Nursery school to grade 4"
"No HS" | "Grade 5, 6, 7, or 8"
"No HS" | "Grade 9"
"No HS" | "Grade 10"
"No HS" | "Grade 11"
"High school graduate" | "Grade 12"
"Some college" | "1 year of college"
"2-year" | "2 years of college"
"4-year" | "4 years of college"
"Post-grad" | "5+ years of college"

```{r}
educ_mapping = list(
    "No HS" = "N/A or no schooling",
    "No HS" = "Nursery school to grade 4",
    "No HS" = "Grade 5, 6, 7, or 8",
    "No HS" = "Grade 9",
    "No HS" = "Grade 10",
    "No HS" = "Grade 11",
    "High school graduate" = "Grade 12",
    "Some college" = "1 year of college",
    "2-year" = "2 years of college",
    "4-year" = "4 years of college",
    "Post-grad" = "5+ years of college"
)
q_educ <- SurveyQuestion$new(name = "educ",
    col_names = c("educ", "EDUC"),
    values_map= educ_mapping)
q_educ$name()
q_educ$col_names()
q_educ$values()
```

@@

```{r}
# create column which conflates spouse present, absent
acs_marst2 <- acs_2018_subset$MARST
levels(acs_marst2) = c(levels(acs_marst2), "Married")
acs_marst2[acs_marst2 =="Married, spouse present"] <- "Married"
acs_marst2[acs_marst2 =="Married, spouse absent"] <- "Married"
acs_marst2 <- droplevels(acs_marst2)  # unused levels not allowed
# add to acs_2018_subset data.frame
acs_2018_subset$marst2 = acs_marst2
```
Now we can create a mapping which maps both "Married" and ""Domestic / civil partnership"
to the the (new) ACS label "Married".^[
Alternatively, we could try to access more ACS variables which account for
same sex marriages and civil unions and combine information across multiple columns, as we did
for ACS variable race2.]

```{r}
marstat_mapping = list(
    "Married" = "Married",
    "Separated" = "Separated",
    "Divorced" = "Divorced",
    "Widowed" = "Widowed",
    "Never married" = "Never married/single",
    "Domestic / civil partnership" = "Married"
)
q_marstat <- SurveyQuestion$new(name = "marstat",
    col_names = c("marstat", "marst2"),
    values_map= marstat_mapping)
q_marstat$name()
q_marstat$col_names()
q_marstat$values()
```

@@

The components of an MRP-KIT mapping are the survey data, the target population data,
and one or more mappings between demographics of interest.  In the previous sections
we assembled the sample and target population datasets and defined the SurveyQuestion
objects for to each demographic, adjusting the datasets accordingly.
Next we create SurveyData objects for each dataset.

A SurveyData object records the contents of a dataset in terms of questions and
responses.  Every column is (potentially) a question, and the responses are the set
of category labels for that column.  (*what about continuous variables?*)
The question and responses are name-value pair lists.
For the question list, the name is the dataframe column name and the value is the description of the variable or the survey question asked.
For the response list, the name is also the dataframe column name and the value is the set of possible responses.

For the CCES, the dataframe column descriptions are lists in the attrbute "var.labels".
As before, we create this list programatically using the `as.list(setNames(tags,values))` one-liner.
```{r}
cces_2018_questions <- as.list(setNames(attr(cces_2018_subset,"var.labels"), names(cces_2018_subset)))
head(cces_2018_questions)
```
For the ACS, we use function `ipumsr::ipums_var_label` to get a short description of the variable
from the metadata in the `acs_2018_subset_tbl` object.
```{r}
acs_2018_var_labels <- c()
for (name in names(acs_2018_subset)) {
    acs_2018_var_labels <- c(acs_2018_var_labels, ipums_var_label(acs_2018_subset_tbl[[name]]))
}
acs_2018_questions <- as.list(setNames(acs_2018_var_labels, names(acs_2018_subset)))
head(acs_2018_questions)
```

A response list is a list of name-value pairs which map the column name to the set of
possible response values.
We create this list by using R's `apply` function.
```{r}
cces_2018_responses <- apply(cces_2018_subset, 2, function(x) levels(as.factor(x)))
head(cces_2018_responses)
acs_2018_responses <- apply(acs_2018_subset, 2, function(x) levels(as.factor(x)))
head(acs_2018_responses)
```

We use these lists to create SurveyData objects for both the CCES and ACS.
```{r}
cces_2018_data <- SurveyData$new(cces_2018_subset, cces_2018_questions, cces_2018_responses)
cces_2018_data
acs_2018_data <- SurveyData$new(acs_2018_subset, acs_2018_questions, acs_2018_responses)
acs_2018_data
```
Now we can create the mapping object from the CCES to the ACS.
The `print` function displays the mappings between the survey data and target population data.
```{r}
cces_acs_mapping <- SurveyMap$new(cces_2018_data, acs_2018_data, q_sex, q_state, q_educ, q_race, q_marstat, q_birthyr)
print(cces_acs_mapping)
```
The `validate` function identifies mapped/unmapped columns and validates the mappings defined by the SurveyQuestion objects
against the levels defined on the datasets.
```{r}
cces_acs_mapping$validate()
```

The `tabulate` function creates the post-stratification table.
```{r}
cces_acs_mapping$tabulate()
```

# MRP Part One:  Multi-level regression

Given the survey data, we fit a model to the survey data, using our favorite R package of choice.

# MRP Part Two:  Post-stratification
